```{r}
library(tidyverse)
library(sf)
library(ggspatial)
library(ggunc)
```

```{r}
data = read_csv("link_scores.csv")
links = read_sf(file.path(Sys.getenv("DATA_PATH"), "data", "links.gpkg"))
```

```{r}
linked = links |>
    right_join(data, by=c("id" = "link_id"))
```

```{r}
with(linked, cor(score, transit_score_delta))
```

```{r}
with(linked, cor(score, transit_score_delta))
```

```{r}
with(linked, cor(score, transit_score_delta, method="spearman"))
```

```{r}
linked |>
    slice_max(score, n = 100) |>
    summarize(median(transit_score_delta))
```

```{r}
linked |>
    ggplot(aes(x=score, y=transit_score_delta)) +
        geom_point(size = 0.5) +
        geom_smooth(method = "lm") +
        xlab("Access to transit score") +
        ylab("Access via transit score") +
        annotate("text", x = Inf, y = Inf, hjust = 1, vjust=1.5, label = sprintf("    Pearson correlation %.2f    \nSpearman correlation %.2f    ", cor(linked$score, linked$transit_score_delta), cor(linked$score, linked$transit_score_delta, method="spearman"))) +
        scale_x_continuous(labels = scales::comma) +
        scale_y_continuous(labels = scales::comma)#, limits = c(0, 7.5e7))

ggsave("paper/figures/comparison.png", width = 6, height = 4, bg="white")
```

```{r}
linked |>
    ggplot(aes(x=rank(score), y=rank(transit_score_delta))) +
        geom_point() +
        geom_smooth(method = "lm") +
        xlab("Access to transit rank") +
        ylab("Access via transit rank") +
        scale_x_continuous(labels = scales::comma)
```

```{r}
ggplot() +
    layer_spatial(st_centroid(st_set_crs(linked, 32119)), aes(color = transit_score_delta), size=0.5) +
    scale_color_binned(breaks = c(0, 100000, 500000, 1000000, 10000000), palette = "reds")
```

## Evaluate links that reduced access (??)

```{r}
reduced_access = linked |>
    filter(transit_score_delta < 0)

View(reduced_access)
```

```{r}
ggplot() +
    annotation_map_tile() +
    layer_spatial(st_centroid(st_set_crs(reduced_access, 32119)))
```

```{r}
linked |>
    st_set_crs(32119)  |>
    write_sf(file.path("/mnt/c/Users/mwbc/OneDrive - University of North Carolina at Chapel Hill/ped_transit_access/", "data", "joined_links.gpkg"))
```

## Baseline


```{r}
blocks = read_sf(file.path(DATA_PATH, "data", "hu.gpkg"))
jobs = read_csv(file.path(DATA_PATH, "data", "nc_wac_S000_JT00_2023.csv.gz"), col_types=cols(w_geocode=col_character()))
```

```{r}
blocks = blocks |>
    # Confirmed same vintage
    left_join(jobs, by=c("GEOID"="w_geocode")) |>
    mutate(C000=replace_na(C000, 0)) |>
    rename(id="GEOID")
```

```{r}
points = st_centroid(blocks) |> st_transform(4326)
```

Get baseline values

```{r}
#write_json(list(minSubgraphSize=unbox(0), stopLinkRadiusMeters=unbox(20)), file.path(DATA_PATH, "networks", "baseline", "config.json"))
base = build_network(file.path(DATA_PATH, "networks", "baseline"), verbose=T, overwrite=T)
acc_base = accessibility(
    r5r_network = base,
    origins = points,
    destinations = points,
    opportunities_colnames = "C000",
    mode = "TRANSIT",
    time_window=120L,
    cutoffs=c(60),
    departure_datetime=as.POSIXct("2025-07-17 07:00:00"),
    progress=T
)

# was: 16658840178
# now: 17102493206
aggregate_access_base = points |>
    left_join(acc_base, by="id") |>
    with(sum(accessibility * H1_001N))

aggregate_access_base
```

## Link 48314

```{r}
link = build_network(file.path(DATA_PATH, "networks", "links", "link_id_48314"), verbose=T, overwrite=T)
acc_link = accessibility(
    r5r_network = link,
    origins = points,
    destinations = points,
    opportunities_colnames = "C000",
    mode = "TRANSIT",
    time_window=120L,
    cutoffs=c(60),
    departure_datetime=as.POSIXct("2025-07-17 07:00:00"),
    progress=T
)

aggregate_access_link = points |>
    left_join(acc_link, by="id") |>
    with(sum(accessibility * H1_001N))

aggregate_access_link
```

```{r}
aggregate_access_link - aggregate_access_base
```

```{r}
acc_link |>
    left_join(acc_base, by="id") |>
    filter(accessibility.x < accessibility.y) |>
    mutate(diff = accessibility.x - accessibility.y)
```

So this one is different, it's not just one location, or if it is it's one with a job center.


```{r}
# so, one location has way lower accessibility and others have marginally lower, so I think that one
# location must have gotten disconnected rather than just a slight time increase due to a rounding error.
orig = points |> filter(id == "370630020152001")

iso_base = isochrone(
    r5r_network = base,
    origins = orig,
    cutoffs = c(60),
    departure_datetime = as.POSIXct("2025-07-17 07:00:00")
)

iso_link = isochrone(
    r5r_network = link,
    origins = orig,
    cutoffs = c(60),
    departure_datetime = as.POSIXct("2025-07-17 07:00:00")
)

problem_pt = points |> filter(str_starts(id, "370630010023000"))

ggplot() +
    annotation_map_tile(alpha = 0.5) +
    annotation_spatial(iso_base, fill="red", alpha = 0.15) +
    annotation_spatial(iso_link, fill="blue", alpha = 0.15) +
    layer_spatial(problem_pt) +
    shadow_spatial(st_buffer(problem_pt, 100))

```

ok so what's going on here is that this block centroid is linked to the parking lot in the baseline, but gets linked to the new link in the scenario, which puts it and its one job just a little out of reach as the new link is a little further from the bus stop on Holloway - and put the rest of the network a little out of reach for it.

```{r}
problem_pt$C000
```

```{r}
ps = map(reduced_access$id, function  (theid) {
    ggplot() +
        annotation_map_tile(alpha=0.5) +
        layer_spatial(reduced_access |> filter(id == theid) |> st_set_crs(32119)) +
        shadow_spatial(reduced_access |> filter(id == theid) |> st_set_crs(32119) |> st_buffer(100)) +
        annotation_spatial(points, color="red") +
        theme_void() +
        ggtitle(theid)
})
```

```{r}
get_acc_link = function (linkid) {
    link = build_network(file.path(DATA_PATH, "networks", "links", paste0("link_id_", linkid)), verbose=T)
    acc_link = accessibility(
        r5r_network = link,
        origins = points,
        destinations = points,
        opportunities_colnames = "C000",
        mode = "TRANSIT",
        time_window=120L,
        cutoffs=c(60),
        departure_datetime=as.POSIXct("2025-07-17 07:00:00"),
        progress=T
    )

    result = acc_link |>
    left_join(acc_base, by="id") |>
    filter(accessibility.x < accessibility.y) |>
    mutate(diff = accessibility.x - accessibility.y)

ggplot() +
    annotation_map_tile(alpha = 0.5, zoomin = 1) +
    layer_spatial(reduced_access |> filter(id ==linkid) |> st_set_crs(32119)) +
    layer_spatial(inner_join(points, result, by = "id")) +
    geom_sf_label(data = inner_join(points, result, by = "id"), aes(label=paste(id, diff)))
}
```

```{r}
get_iso_link = function (linkid, origin) {
    link = build_network(file.path(DATA_PATH, "networks", "links", paste0("link_id_", linkid)), verbose=T)

    pt = points[points$id == origin,]

    iso_link = isochrone(
        r5r_network = link,
        origins = pt,
        mode = "TRANSIT",
        time_window=120L,
        cutoffs=c(60),
        departure_datetime=as.POSIXct("2025-07-17 07:00:00"),
        zoom = 11
    )

    iso_base = isochrone(
        r5r_network = base,
        origins = pt,
        mode = "TRANSIT",
        time_window=120L,
        cutoffs=c(60),
        departure_datetime=as.POSIXct("2025-07-17 07:00:00"),
        zoom = 11
    )

    ggplot() +
        annotation_map_tile(alpha = 0.5, zoomin = 1) +
        layer_spatial(iso_base, fill = "blue", alpha = 0.25) +
        layer_spatial(iso_link, fill = "red", alpha = 0.25) +
        layer_spatial(pt) +
            layer_spatial(reduced_access |> filter(id ==linkid) |> st_set_crs(32119))
    }
```

```{r}
ttm_base = travel_time_matrix(
    r5r_network = base,
    origins = points,
    destinations = points,
    mode = "TRANSIT",
    time_window=120L,
    departure_datetime=as.POSIXct("2025-07-17 07:00:00")
)

get_ttm_link = function (linkid) {
    link = build_network(file.path(DATA_PATH, "networks", "links", paste0("link_id_", linkid)), verbose=T)

    ttm_link = travel_time_matrix(
        r5r_network = link,
        origins = points,
        destinations = points,
        mode = "TRANSIT",
        time_window=120L,
        departure_datetime=as.POSIXct("2025-07-17 07:00:00"),
        progress = TRUE
    )

    result = ttm_link |>
        left_join(ttm_base, join_by(from_id, to_id), suffix = c(".link", ".base")) |>
        left_join(select(points, id, H1_001N), join_by(from_id == id)) |>
        left_join(select(points, id, C000), join_by(to_id == id))

    print(sprintf("Largest difference: %d", max(result$travel_time_p50.link - result$travel_time_p50.base)))

    return(result)
}
```

```{r}
map_affected_trips = function (linkid, ttm) {
    dat = ttm |>
        filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60) |>
        rowwise() |>
        mutate(geometry = list(st_linestring(matrix(c(st_coordinates(geom.x), st_coordinates(geom.y)), nrow=2, byrow=T)))) |>
        st_drop_geometry() |>
        select(geometry) |>
        st_as_sf() |>
        st_set_crs(4326)

    ggplot() +
        annotation_map_tile(alpha = 0.5, zoomin = 1) +
        annotation_spatial(points) +
        layer_spatial(reduced_access |> filter(id ==linkid) |> st_set_crs(32119)) +
        layer_spatial(dat, color="red")
    

}
```


```{r}
link_net_to_sf = function (linkid) {
    link = build_network(file.path(DATA_PATH, "networks", "links", paste0("link_id_", linkid)), verbose=T)
    street_network_to_sf(link)$edges
}
```

## Link 862
```{r}
ps[[1]]
```

```{r}
ttm = get_ttm_link(862)

ttm |>
    filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60)
```

```{r}
map_affected_trips(862, ttm)
```

This is pretty obviously due to linking changes.

## Link 6200 (-156)

```{r}
ps[[2]]
```

```{r}
get_acc_link(6200)
```

```{r}
ttm6200 = get_ttm_link(6200)

ttm6200 |>
    filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60)
```

```{r}
get_iso_link(6200, 371830537301002)
```

This has to be rounding.

## Link 13267 (-18)

```{r}
ps[[3]]
```

```{r}
get_acc_link(13267)
```

```{r}
get_ttm_link(13267) |>
    filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60)
```

Gotta be rounding.

## Link 22763

```{r}
ps[[4]]
```

```{r}
get_acc_link(22763)
```

```{r}
get_ttm_link(22763) |>
    filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60)
```

This one forces a double-back by linking to the new link instead of to the parking aisle.

## Link 24140

```{r}
ps[[5]]
```

```{r}
ttm = get_ttm_link(24140)

ttm |>
    filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60)
```



```{r}
map_affected_trips(24140, ttm)
```

This one is just that these trips are probably walk trips that became marginally longer due to snapping.

## Link 27541 (-39215)

```{r}
ps[[6]]
```

```{r}
ttm = get_ttm_link(27541)

ttm |>
    filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60)
```

```{r}
map_affected_trips(27541, ttm)
```

```{r}
get_acc_link(27541)
```

Not exactly sure here because the affected trips are mostly in opposite direction, but it's gotta be some kind of rounding error.


## Link 48314

```{r}
ps[[7]]
```

Snapping changed, no surprise here.

## Link 58495 (-24)

```{r}
ps[[8]]
```

```{r}
ttm = get_ttm_link(58495)

ttm |>
    filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60)
```

```{r}
map_affected_trips(58495, ttm)
```

Gotta be rounding.

## Link 59463

```{r}
ps[[9]]
```

```{r}
ttm = get_ttm_link(59463)

ttm |>
    filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60)
```

```{r}
map_affected_trips(59463, ttm)
```

This must be snapping.

## Link 5969
6

```{r}
ps[[10]]
```

```{r}
get_acc_link(59696)
```

```{r}
get_iso_link(59696, "371830535232012")
```

```{r}
ggplot() +
    annotation_map_tile(alpha = 0.5, zoomin = 1) +
    layer_spatial(reduced_access |> filter(id == 69007) |> st_set_crs(32119)) +
    layer_spatial(inner_join(points, link59696, by = "id"))
```

```{r}
ttm59696 = get_ttm_link(59696)
```

My conclusion here is that this one is another false positive - I'm not exactly sure what is happening, I suspect it's rouding

```{r}
street_network_to_sf(base)$edges |>
    write_sf("base.gpkg")
```

```{r}
link_net_to_sf(59696) |>
    write_sf("59696.gpkg")
```

IDK this has to be rounding.

## Link 69007

```{r}
ps[[11]]
```

```{r}
link69007 = get_acc_link(69007)
```

OK, so this one is just that snapping makes the route marginally longer, and there are a handful of places where that marginally-longer matters.