```{r}
library(tidyverse)
library(sf)
library(ggspatial)
library(ggunc)
library(gtfstools)
library(ggbreak)
options(java.parameters="-Xmx8G", r5r.r5jar="../r5/build/libs/r5-v4.1.0-124-g996dd94-all.jar")
devtools::load_all("../r5r/r-package")

DATA_PATH = Sys.getenv("DATA_PATH")
```

```{r}
data = read_csv("link_scores.csv")
links = read_sf(file.path(Sys.getenv("DATA_PATH"), "data", "links.gpkg"))
```

```{r}
linked = links |>
    right_join(data, by=c("id" = "link_id"))
```

```{r}
with(linked, cor(score, transit_score_delta))
```

```{r}
with(linked, cor(score, transit_score_delta, method="spearman"))
```

```{r}
linked |>
    slice_max(score, n = 100) |>
    summarize(median(transit_score_delta))
```

```{r}
pfull = linked |>
    ggplot(aes(x=score, y=to_broken_vals(transit_score_delta))) +
        geom_point(size = 0.5) +
        geom_smooth(method = "lm") +
        xlab("Access to transit score") +
        ylab("Access via transit score") +
        annotate("text", x = Inf, y = Inf, hjust = 1, vjust=1.5, label = sprintf("    Pearson correlation %.2f    \nSpearman correlation %.2f    ", cor(linked$score, linked$transit_score_delta), cor(linked$score, linked$transit_score_delta, method="spearman"))) +
        scale_x_continuous(labels = scales::comma) +
        scale_y_continuous(labels = scales::comma)#, limits = c(0, 7.5e7))


ggsave("paper/figures/comparison.png", width = 6, height = 4, bg="white")
```

The spearman correlation is highly statistically significant.

```{r}
lm(rank(transit_score_delta) ~ rank(score), linked) |> summary()
```

```{r}
linked |>
    ggplot(aes(x=rank(score), y=rank(transit_score_delta))) +
        geom_point() +
        geom_smooth(method = "lm") +
        xlab("Access to transit rank") +
        ylab("Access via transit rank") +
        scale_x_continuous(labels = scales::comma)
```

```{r}
median(linked[linked$rank <= 200,]$transit_score_delta)
```

```{r}
ggplot() +
    layer_spatial(st_centroid(st_set_crs(linked, 32119)), aes(color = transit_score_delta), size=0.5) +
    scale_color_binned(breaks = c(0, 100000, 500000, 1000000, 10000000), palette = "reds")
```

## Evaluate links that reduced access (??)

```{r}
reduced_access = linked |>
    filter(transit_score_delta < 0)

reduced_access |>
    st_drop_geometry() |>
    select(id, transit_score_delta)
```

## Baseline


```{r}
blocks = read_sf(file.path(DATA_PATH, "data", "hu.gpkg"))
jobs = read_csv(file.path(DATA_PATH, "data", "nc_wac_S000_JT00_2023.csv.gz"), col_types=cols(w_geocode=col_character()))
```

```{r}
blocks = blocks |>
    # Confirmed same vintage
    left_join(jobs, by=c("GEOID"="w_geocode")) |>
    mutate(C000=replace_na(C000, 0)) |>
    rename(id="GEOID")
```

```{r}
points = st_centroid(blocks) |> st_transform(4326)
```

Get baseline values

```{r}
#write_json(list(minSubgraphSize=unbox(0), stopLinkRadiusMeters=unbox(20)), file.path(DATA_PATH, "networks", "baseline", "config.json"))
base = build_network(file.path(DATA_PATH, "networks", "baseline"), verbose=T, overwrite=T)
acc_base = accessibility(
    r5r_network = base,
    origins = points,
    destinations = points,
    opportunities_colnames = "C000",
    mode = "TRANSIT",
    time_window=120L,
    cutoffs=c(60),
    departure_datetime=as.POSIXct("2025-07-17 07:00:00"),
    progress=T
)

# was: 16658840178
# now: 17102493206
# now: 16952351630
aggregate_access_base = points |>
    left_join(acc_base, by="id") |>
    with(sum(accessibility * H1_001N))

aggregate_access_base
```

## Functions for evaluating link changes

### Plot them all

```{r}
ps = map(reduced_access$id, function  (theid) {
    ggplot() +
        annotation_map_tile(alpha=0.5) +
        layer_spatial(reduced_access |> filter(id == theid) |> st_set_crs(32119)) +
        shadow_spatial(reduced_access |> filter(id == theid) |> st_set_crs(32119) |> st_buffer(100)) +
        annotation_spatial(points, color="red") +
        theme_void() +
        ggtitle(sprintf("Link %d, access change %d", theid, reduced_access[reduced_access$id == theid, "transit_score_delta"][[1]]))
})
```

```{r}
get_acc_link = function (linkid) {
    link = build_network(file.path(DATA_PATH, "networks", "links", paste0("link_id_", linkid)), verbose=T)
    acc_link = accessibility(
        r5r_network = link,
        origins = points,
        destinations = points,
        opportunities_colnames = "C000",
        mode = "TRANSIT",
        time_window=120L,
        cutoffs=c(60),
        departure_datetime=as.POSIXct("2025-07-17 07:00:00"),
        progress=T
    )

    result = acc_link |>
    left_join(acc_base, by="id") |>
    filter(accessibility.x < accessibility.y) |>
    mutate(diff = accessibility.x - accessibility.y)

ggplot() +
    annotation_map_tile(alpha = 0.5, zoomin = 1) +
    layer_spatial(reduced_access |> filter(id ==linkid) |> st_set_crs(32119)) +
    layer_spatial(inner_join(points, result, by = "id")) +
    geom_sf_label(data = inner_join(points, result, by = "id"), aes(label=paste(id, diff)))
}
```

```{r}
get_iso_link = function (linkid, origin, cutoff = 60) {
    link = build_network(file.path(DATA_PATH, "networks", "links", paste0("link_id_", linkid)), verbose=T)

    pt = points[points$id == origin,]

    iso_link = isochrone(
        r5r_network = link,
        origins = pt,
        mode = "TRANSIT",
        time_window=120L,
        cutoffs=c(cutoff),
        departure_datetime=as.POSIXct("2025-07-17 07:00:00"),
        zoom = 11
    )

    iso_base = isochrone(
        r5r_network = base,
        origins = pt,
        mode = "TRANSIT",
        time_window=120L,
        cutoffs=c(cutoff),
        departure_datetime=as.POSIXct("2025-07-17 07:00:00"),
        zoom = 11
    )

    ggplot() +
        annotation_map_tile(alpha = 0.5, zoomin = 1) +
        layer_spatial(iso_base, fill = "blue", alpha = 0.25) +
        layer_spatial(iso_link, fill = "red", alpha = 0.25) +
        layer_spatial(pt) +
            layer_spatial(reduced_access |> filter(id ==linkid) |> st_set_crs(32119))
    }
```

```{r}
ttm_base = travel_time_matrix(
    r5r_network = base,
    origins = points,
    destinations = points,
    mode = "TRANSIT",
    time_window=120L,
    departure_datetime=as.POSIXct("2025-07-17 07:00:00")
)

get_ttm_link = function (linkid, dir="links") {
    link = build_network(file.path(DATA_PATH, "networks", dir, paste0("link_id_", linkid)), verbose=T)

    ttm_link = travel_time_matrix(
        r5r_network = link,
        origins = points,
        destinations = points,
        mode = "TRANSIT",
        time_window=120L,
        departure_datetime=as.POSIXct("2025-07-17 07:00:00"),
        progress = TRUE
    )

    result = ttm_link |>
        left_join(ttm_base, join_by(from_id, to_id), suffix = c(".link", ".base")) |>
        left_join(select(points, id, H1_001N), join_by(from_id == id)) |>
        left_join(select(points, id, C000), join_by(to_id == id))

    print(sprintf("Largest difference: %d", max(result$travel_time_p50.link - result$travel_time_p50.base, na.rm=T)))

    return(result)
}
```

```{r}
map_affected_trips = function (linkid, ttm) {
    dat = ttm |>
        filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60) |>
        rowwise() |>
        mutate(geometry = list(st_linestring(matrix(c(st_coordinates(geom.x), st_coordinates(geom.y)), nrow=2, byrow=T)))) |>
        st_drop_geometry() |>
        select(geometry) |>
        st_as_sf() |>
        st_set_crs(4326)

    ggplot() +
        annotation_map_tile(alpha = 0.5, zoomin = 1) +
        annotation_spatial(points, size=0.1) +
        layer_spatial(reduced_access |> filter(id ==linkid) |> st_set_crs(32119)) +
        layer_spatial(dat, color="red")
    

}
```


```{r}
link_net_to_sf = function (linkid) {
    link = build_network(file.path(DATA_PATH, "networks", "links", paste0("link_id_", linkid)), verbose=T)
    street_network_to_sf(link)$edges
}
```

## Link 5294

Gotta be rounding.

```{r}
ps[[1]]
```

```{r}
ttm = get_ttm_link(5294)

ttm |> filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60) |>
    st_drop_geometry() |>
    select(ends_with("id"), starts_with("travel"))
```

```{r}
map_affected_trips(5294, ttm)
```

## Link 8525

This one's pretty clearly a linking change.

```{r}
ps[[2]]
```

```{r}
ttm = get_ttm_link(8525)

ttm |> filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60) |>
    st_drop_geometry() |>
    select(ends_with("id"), starts_with("travel"))
```

```{r}
map_affected_trips(8525, ttm)
```

## Link 9630

Looks like linking, slightly longer path.

This is a good example of the issues with network construction: https://www.openstreetmap.org/search?query=mingocrest+dr&zoom=18&minlon=-79.03669595718385&minlat=35.92202502675714&maxlon=-79.03111159801485&maxlat=35.92520056770574#map=18/35.792199/-78.497639

Sidewalk opposite not connected to cul de sac so this link is useless.

```{r}
ps[[3]]
```

```{r}
ttm = get_ttm_link(9630)

ttm |> filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60) |>
    st_drop_geometry() |>
    select(ends_with("id"), starts_with("travel"))
```

```{r}
map_affected_trips(9630, ttm)
```

## Link 10896

This changes how this stop is getting linked - it is now linked to a sidewalk fragment.

```{r}
ps[[4]]
```

```{r}
ttm = get_ttm_link(10896)

ttm |> filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60) |>
    st_drop_geometry() |>
    select(ends_with("id"), starts_with("travel"))
```

```{r}
map_affected_trips(10896, ttm)
```

## Link 21914

Looks like rounding.

```{r}
ps[[5]]
```

```{r}
ttm = get_ttm_link(21914)

ttm |> filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60) |>
    st_drop_geometry() |>
    select(ends_with("id"), starts_with("travel"))
```

```{r}
map_affected_trips(21914, ttm)
```

## Link 40463

Looks like slight linking differences.

```{r}
ps[[6]]
```

```{r}
ttm = get_ttm_link(40463)

ttm |> filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60) |>
    st_drop_geometry() |>
    select(ends_with("id"), starts_with("travel"))
```

```{r}
map_affected_trips(40463, ttm)
```

## Link 40676

Probably rounding.

```{r}
ps[[7]]
```

```{r}
ttm = get_ttm_link(40676)

ttm |> filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60) |>
    st_drop_geometry() |>
    select(ends_with("id"), starts_with("travel"))
```

```{r}
map_affected_trips(40676, ttm)
```

## Link 41591

Gotta be rounding.

```{r}
ps[[8]]
```

```{r}
ttm = get_ttm_link(41591)

ttm |> filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60) |>
    st_drop_geometry() |>
    select(ends_with("id"), starts_with("travel"))
```

```{r}
map_affected_trips(41591, ttm)
```

## Link 41745

Slight linking differences.

```{r}
ps[[9]]
```

```{r}
ttm = get_ttm_link(41745)

ttm |> filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60) |>
    st_drop_geometry() |>
    select(ends_with("id"), starts_with("travel"))
```

```{r}
map_affected_trips(41745, ttm)
```

## Link 43000

```{r}
ps[[10]]
```

```{r}
ttm = get_ttm_link(43000)

ttm |> filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60) |>
    st_drop_geometry() |>
    select(ends_with("id"), starts_with("travel"))
```

```{r}
map_affected_trips(43000, ttm)
```

## Link 44988

I think this is a linking change for the bus stop.

```{r}
ps[[11]]
```

```{r}
ttm = get_ttm_link(44988)

ttm |> filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60) |>
    st_drop_geometry() |>
    select(ends_with("id"), starts_with("travel"))
```

```{r}
map_affected_trips(44988, ttm)
```

## Link 49528

Small linking change.

```{r}
ps[[12]]
```

```{r}
ttm = get_ttm_link(49528)

ttm |> filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60) |>
    st_drop_geometry() |>
    select(ends_with("id"), starts_with("travel"))
```

```{r}
map_affected_trips(49528, ttm)
```

## Link 55593

So this is a big change, there's a bus linking change here that increases the walking distance between the 23L and the 8 and causes you to miss the transfer.

```{r}
ps[[13]]
```

```{r}
ttm = get_ttm_link(55593)

ttm |> filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60) |>
    st_drop_geometry() |>
    select(ends_with("id"), starts_with("travel"))
```

```{r}
map_affected_trips(55593, ttm)
```

```{r}
get_iso_link(55593, "371830537151010")
```

```{r}
ggplot() +
    annotation_map_tile() +
    layer_spatial(
        blocks |>
            inner_join(
        ttm |>
        filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60),
        by = c("id" = "from_id"))
    )
```

```{r}
gora = read_gtfs(file.path(DATA_PATH, "data", "gtfs", "goraleigh.zip")) |>
    filter_by_service_id(c("1", "5"))

# transfer from 23L
with(gora,
    trips |>
    filter(route_id == "23L" & direction_id == 0) |>
    inner_join(stop_times, join_by(trip_id)) |>
    filter(stop_id == "2790") |> # E Millbrook at Six Forks
    pull(departure_time) |>
    sort()
)
```

```{r}
# Transfer to 8
with(gora,
    trips |>
    filter(route_id == "8" & direction_id == 0) |>
    inner_join(stop_times, join_by(trip_id)) |>
    filter(stop_id == "3009") |> # Six Forks at Millbrook
    pull(departure_time) |>
    sort()
)
```

## Link 57672

Definitely linking changes

```{r}
ps[[14]]
```

```{r}
ttm = get_ttm_link(57672)

ttm |> filter(travel_time_p50.link >= 60 & travel_time_p50.base < 60) |>
    st_drop_geometry() |>
    select(ends_with("id"), starts_with("travel"))
```

```{r}
map_affected_trips(57672, ttm)
```